---
title: "Basics in R"
author: "Denise J. Roth"
date: "`r Sys.Date()`"
output: html_document
---
# Basics in R


## What are functions in R? 

A function in R is a block of code that performs a specific task when called. They typical take input parameters, perform operations on them, and then return a result based on these operations. They encapsulate commonly used operations into a single unit, making one's own work more transparent and reproducible. 
R already comes with a lot of built-in functions. 

Let's look at an example:


```{r Example of built-in function}
# Creating a vector of numbers
numbers <- c(2, 4, 6, 8, 10)

# Calculate the mean of the numbers using the mean()
average <- mean(numbers)

# Print the result
print(average)
```

The flexibility of R allows us to also write our own functions for very specific tasks and operations. 

What could this look?

```{r Writing a function}
# Defining the function
addition <- function(x, y) {
  sum <- x + y
  return(sum)
}

# Call the function with arguments 3 and 5
result <- addition(3, 5)

# Print the result
print(result)
```

## What is the meaning of the "->"-operator?

In R, the "->" symbol is used for assigning values to variables. It's like saying "put this value into that box." So when you see something like x -> y, it means the value of x is stored in the variable y.

R is an _object-oriented_ programming language - in simple terms, it means that everything in R is treated as an object. An object is like a container that holds data and functions that can work with that data. So when you're working in R, you're usually creating, manipulating, or using objects.

For example, when you create a variable like x <- 5, x is an object that holds the value 5. And when you perform operations or functions on x, you're essentially interacting with that object.

In our environment, we can see the objects to which we have assigned any sort of value etc.


```{r Objects}
random_object <- 27

print(random_object)
```

## Packages in R

In R, a package is like a toolbox filled with useful tools for specific tasks. Each package contains a collection of functions, datasets, and other resources designed to help you tackle certain problems more efficiently.

Packages make it easier to extend R's capabilities beyond its built-in functions. They're created and maintained by members of the R community, and you can easily install them using the ```install.packages()``` function.


```{r Installing packages}
install.packages("tidyverse")
```


Let's try using some functions that is found in the ```tidyverse``` package!


```{r}

# Defining an example data frame
example_data <- tibble(
  name = c("Alice", "Bob", "Charlie"),
  age = c(25, 30, 35)
)

# Attemptong to filtering data and keeping only the cases older than 30
plus_30s <- example_data |> 
  filter(age > 30)
```

*Important*: If we want to use a package and its functions/datasets/resources, we need to use the ```library()``` function to load the package. Otherwise, R will not know what we are referring to. 


